\Large \bf \Toolname: Precise False Sharing Detection for COTS Binaries

### 2014-11-05 HotOS'15/paper.tex
Renamed design to methodology, used a bit of the same style as engler.. mwahahaha. Strengthening the idea of beliefs about code, as I think that accurately describes what is being done.

False-sharing is a performance-degrading problem that limits the scalability of multithreaded applications
\cite{ImpactOfFalseSharing}. This problem is increasingly important in the multi-core era. The level of
concurrency in applications is growing to match the additional processing power provided by increasing
core counts on modern processors. Despite this rise in importance, detecting and diagnosing false
sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
sharing is often indistinguishable from true sharing (including race conditions). Also, diagnosing the cause
of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.

In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
in multi-threaded, commercial off-the-shelf (COTS) binaries. The key insights of our approach is that
it iteratively discovers and specializes instrumentation to detect false sharing at increasingly finer
granularities. Our work addresses three limitations of prior work:
\begin{itemize}
\item They introduce extra cache coherency traffic, which limits scalability and hides performance
problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.

@@

\end{itemize}

\paragraph{Contributions}
This paper makes the following contributions:
\begin{itemize}
\item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
false sharing  in multi-threaded, CTOS binaries. We believe that operating directly on binaries is
important because it allows one to verify observed performance problems using the same binaries
that are running in a production environment.

\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
efficient determination of memory access behavior using shadow memory. Unlike existing shadow
memory systems, proxy memory is not treated as a mutable meta-data store. This is important
for false sharing detection as it avoids the extra serialization and coherence traffic that is introduced
by dynamically manipulating shadow state.

@@

% --------------------------------------------------------------------------------------
\section{Design}\label{sec:design}

% --------------------------------------------------------------------------------------
\section{Implementation}\label{sec:implementation}
\Toolname{} is implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
\cite{AddressWatchpoints} framework.

\section{Status}\label{sec:status}

\section{Evaluation}\label{sec:evaluation}

>>>>

False-sharing is a performance-degrading problem that limits the scalability of multithreaded applications
\cite{ImpactOfFalseSharing}. This problem is increasingly important in the multi-core era. The level of
concurrency in applications is growing to match the additional processing power provided by increasing
core counts on modern processors. Despite this rise in importance, detecting and diagnosing false
sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
sharing is often indistinguishable from true sharing (including data races). Also, diagnosing the cause
of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.

In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
in multi-threaded applications. \Toolname{} uses dynamic binary translation (DBT) and shadow memory
to detect false sharing in commercial off-the-shelf (COTS) binaries.  The key insight of our
approach is that accesses to shadow memory can be used as a proxy for accesses to actual memory. By 
detecting contention on proxied accesses, we can establish beliefs about what code causes false sharing. These
beliefs are strenghtened and eventually verified by specializing code to proxy memory accesses at increasingly
finer granularities.

Our work addresses three limitations of prior work:
\begin{enumerate}
\item They introduce extra cache coherency traffic, which limits scalability and hides performance
problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.

@@

\end{enumerate}

\paragraph{Contributions}
This paper makes the following significant contributions:
\begin{itemize}
\item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
false sharing  in multi-threaded, CTOS binaries. We believe that operating directly on binaries is
important because it allows one to verify observed performance problems using the same binaries
that are running in a production environment.

\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
efficient determination of memory access behavior using shadow memory. Unlike existing shadow
memory systems, proxy memory is not treated as a mutable meta-data store. This avoids extra
serialization and coherence traffic introduced by prior work when dynamically manipulating shadow state.

@@

% --------------------------------------------------------------------------------------
\section{Methodology}\label{sec:methodology}

This section introduces our approach to finding false sharing bugs. The goal of our approach is extract beliefs about
where false sharing might be, then iteratively strenghten those beliefs, either verifying or discarding them in the process.

We begin with the same intuition as SHERIFF: if false sharing occurs on a single cache line, then that implies that two
threads are simultaneously accessing the same page of memory \cite{SHERIFF}. Therefore, by detecting contention on
a single page, we can establish a belief that false sharing might occur on the contended page.

Unfortunately, a page-granularity detection of false sharing is na{\"i}ve: it provides no guidance as to the location of the
contended cache line, if any. Our insight is that we can narrow down on instances of false sharing by varying the
granularity of the mechanism used to detect thread contention. That is, we can start with a coarse-grained but
innacurate approach that tells us what code might cause false sharing, thus establishing an initial belief set. Then, we
can strengthen or discard beliefs by repeating this same process, but using finer-grained detection mechanisms that
apply only to code thought to contain false sharing.

The remainder of this section explains how we vary the detection granularity (\Secref{proxymem}), how we use sampling
to detect contending accesses to shared memory (\Secref{heapsample}), how we reduce false negatives with our
sampling-based approach (\Secref{falsenegs}), and finally how we rank discovered instances of false sharing (\Secref{ranking})

\subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}

\subsection{Heap Sampling}\label{sec:heapsample}

\subsection{Reducing False Negatives}\label{sec:falsenegs}

One challenge with sampling-based approaches is their ability to get ``unlucky" and either not sample
the right thing, or sample the right thing at the wrong time. In our case, this results in false-negatives:
we might report that no false sharing is detected in some code, but this does not imply the absence of
false sharing.

To address this challenge, we propagate existing beliefs about potential false sharing on a single instruction
to nearby instructions, as well as to instructions that operate on data of the same type. The intuition is
two-fold. First, we assume locality of accesses to shared data: if we observe a shared access on one instruction,
then nearby instructions probably access shared data as well. Second, a belief of false sharing is a belief of
sharing: if we observe an object of a particular type being concurrently accessed in one place, then it might
imply that concurrent accesses can also happen in other places.

\newcommand\TypeId{$TypeId$}

We perform two separate analyses to extend belief sets: \begin{description}
\item[Type Analysis] \hfill \\
This analysis instruments all code and identifies what types of objects are accessed by what code. Object types are loosely
defined as \TypeIdPair{} pairs, where $PC$ is the return address of an allocation function, and $size$ is the number of bytes
being allocated. To each \TypeIdPair pair, we associate a unique \TypeId{}. We taint \cite{AddressWatchpoints} allocated
addresses with their \TypeId{}, and propagate this taint to code accessing typed data.

\item[Concurrency Analysis] \hfill \\
This analysis instruments all code and identifies what functions can execute concurrently. Each instrumented function
is associated with a core set and a threadset \cite{DrContention}. When executed, the coreset is updated to include the
identifier of the CPU core executing the function. Similarly, the function's threadset is updated to include an approximation
of the thread identifier ($SP >> 13$). We anlayse the coresets and threadsets of each function to determine whether or
not a function can execute concurrently (across CPU cores, or in two threads on the same core).

\end{description}

We combine the type and concurrency analyses to propagate beliefs in the following way: if sampling detects potential
false sharing on a single instruction operating on a specific type of object, then we mark all instructions in the intersection
of those identified to operate on the same type with the set of concurrent functions.

\subsection{Ranking}\label{sec:ranking}

Because we sample data of all types, we might find instances of false sharing that have no significant impact on performance.
Therefore, we need a means of estimating the performance impact of false sharing so that we can rank detected instances of
false sharing from most to least important. We accomplish this as part of a separate analysis step. This step
instruments all code to count how often each basic block is executed. We correlate this information with detected
instances of false sharing to infer cases where one or both threads are executing hot code and contending on a single
cache line. In the former case, we apply a variant of the cold code hypothesis \cite{LiteRace} to rank cases of cold
code ``pessimizing" hot code more highly.

%At a high-level, \Toolname{} follows the same principle as Plastic \cite{Plastic} and SHERIFF \cite{SHERIFF}: it detects
%false sharing at a coarse granularity and then verifies detected cases at a finer granularity. Unlike other approaches,
%\Toolname{} repeats this process by incrementally refining the detection granularity, and by selectively instrumenting only
%that code that appears to trigger false sharing. Whereas Plastic and SHERIFF use page-granularity detection, \Toolname{}
%is able to vary the granularity over time using a \emph{proxy memory} (\Secref{proxymem}).

% --------------------------------------------------------------------------------------
\section{Status}\label{sec:status}
\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
\cite{AddressWatchpoints} framework.

The following has been implemented:
% TODO...

% --------------------------------------------------------------------------------------
\section{Evaluation}\label{sec:evaluation}
